### 2022-06-15 Seminar/preamble.tex
Adding seminar files

%load any additional packages

\usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc} % so we can input characters with accents (e.g. ő)

\usepackage{statsbeamer}
\usepackage{multirow,booktabs}%pacotes para tabelas
\usepackage{subcaption}
\usepackage[font=footnotesize,skip=2pt]{caption}
%\captionsetup[table]{skip=2pt}
\usepackage{ragged2e}
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\tikzset{%
  >={Latex[width=2mm,length=2mm]},
  % Specifications for style of nodes:
            base/.style = {rectangle, rounded corners, draw=black,
                           minimum width=4cm, minimum height=1cm,
                           text centered, font=\sffamily},
  activityStarts/.style = {base, fill=oxfordblue!30},
       startstop/.style = {base, fill=oxfordblue!30},
    activityRuns/.style = {base, fill=oxfordblue!30},
         process/.style = {base, minimum width=2.5cm, fill=orange!15},
}

\usepackage{graphicx} % ease graphics management
\graphicspath{{figs/}} % define folder with images
\usefonttheme{serif} % change font to allow \textbf{}
\usepackage{charter} % Nicer fonts

\usepackage{cancel}
\usepackage{amsmath,amsthm,amssymb} % for math equations

\usepackage[numbers]{natbib} % richer citation
\usepackage{breakcites} % avoid overfull hbox for long cites

\usepackage{algorithm,algorithmic}                                          % Pseudo-code environment
\usepackage{amsmath,amsfonts,amsthm,amssymb,bm,mathdots,mathtools,bigints}  % Math packages

\usepackage{caption}                        % Tools for customizing the \caption{} command
\usepackage{subcaption}						% Environment to create subcaptions

\captionsetup{singlelinecheck = false, format= hang, justification=raggedright, font=footnotesize, labelsep=space} % ABNT Format for Caption